Search for Spatially Extended \fermi-LAT Sources Using Two Years of Data

### 2011-11-10 extended_catalog/paper/extended_catalog.tex
updating draft

The current generation of air Cherenkov detectors have
made it apparent that many sources can be spatially resolved
at even higher energies.
Most
prominent was a survey of the Galactic plane using the High Energy
Stereoscopic System (H.E.S.S) which reported 14 spatially extended
sources with extensions varying from $\sim0\fdg1$ to $\sim0\fdg25$
\citep{hess_plane_survey}.  In fact, within our Galaxy only very few
sources detected at \tev energies, most notably the $\gamma$-ray binaries
LS\,5039 \citep{HESSLS5039}, LS I+61$-$303 \citep{MAGICLSI, VERITASLSI},
HESS\,J0632+057 \citep{HESS0632}, and the Crab nebula \citep{crab_weekes},
have no detectable extension.  High-energy $\gamma$-rays from
these sources are produced by the decay of $\pi^0$s produced by hadronic
interactions with interstellar matter and by relativistic electrons
due to Inverse Compton (IC) scattering and bremsstrahlung radiation.  
It is plausible that the \gev and
\tev emission from these sources originates from the same population of
high-energy particles and so at least some of these sources should be
detected at \gev energies.  Studying 
these \tev sources at \gev energies would help to
determine the emission mechanisms producing these high energy photons.

@@

\newtext{
A new maximum-likelihood analysis tool has been developed to address the
unique requirements for studying spatially extended sources with the LAT.
It works by maximizing the Poisson likelihood to find the observing counts
given a parametrized spatial and spectral model of the sky.  
Space is
divided into cubes of position and energy using the healpix representation of
the sky \citep{healpix} and the likelihood is maximized over all bins in
a region.
}
The extension of a source can be modeled by a geometric shape
(e.g. a disk or a two-dimensional Gaussian) and the position, extension,
and spectrum of the source can be simultaneously fit.

This type of analysis is not feasible using the standard LAT likelihood
analysis tool \gtlike\footnote{\gtlike is distributed publicly by the
FSSC.} because it can only fit the spectral parameters of the model
unless a more sophisticated iterative procedure is used to also test
various source morphologies.  We note that \gtlike has been used in the
past in several studies of source extension in the LAT Collaboration
\citep{lmc,smc,w28,w51c}.  In these studies, a profile based upon a
set of \gtlike maximum likelihood fits at fixed extensions was used
to build a profile of the likelihood as a function of extension.
\newtext{
The \gtlike likelihood profile approach has been show to correctly
reproduce the extension of simulated extended sources assuming that the
true position is known \citep{francesco_2011}.  But it is not optimal
because the position, extension, and spectrum of the source must be
simultaneously fit to find the best fit parameters and to maximize the
statistical significance of the detection.  Furthermore, because the \gtlike
approach is computationally intensive, no large-scale Monte Carlo
simulations have been run to calculate its false detection rate.
}

@@

In \pointlike, it is assumed that the spatial and spectral model
of an extended source are separable, i.e. that the source model
$M(l,b,E)=S(l,b)\times X(E)$ where $S(l,b)$ is the spatial distribution
and $X(E)$ is the spectral distribution.  To fit an extended source,
\pointlike convolves the extended source shape with the PSF (as a function
of energy) and uses the \minuit library \citep{minuit_documentation}
to maximize the likelihood by simultaneously varying the position,
extension, and spectrum of the source.  As will be described in
Section~\ref{monte_carlo_validation}, simultaneously fitting the
position, extension, and spectrum is important to maximize
the statistical significance of the detection of extension of a source.
To avoid projection effects, the longitude and latitude 
of the source are
not directly fit but instead the displacement of the source in a rotated
reference frame.

@@

For extended sources with an assumed radially-symmetric shape,
we optimized the calculation by performing one
of the integrals analytically.
The expected photon 
distribution can be written as
\begin{equation}
  \text{PDF}(\vec r) = \int  \text{PSF}(|\vec r - \vec r'|)I_\text{src}(\vec r') r' dr' d\phi'
\end{equation}
where $I_\text{src}(\vec r')$ is the spatial distribution of the
source.
\newtext{
The LAT's PSF is currently parameterized 
in the Pass~7\_V6 (P7\_V6) Source Instrument
Response Function \citep[IRFs,][]{lat_on_orbit_psf} by a King function \citep{king_function}:
}
\begin{equation}
  \text{PSF}(r) = 
  \frac{1}{2\pi\sigma^2}
  \left(1-\frac{1}{\gamma}\right)
  \left(1+\frac{u}{\gamma}\right)^{-\gamma},
\end{equation}
where $u=(r/\sigma)^2/2$ and $\sigma$ and $\gamma$ are free parameters
\citep{matthew_kerr_thesis}.  For radially-symmetric extended sources,
the angular part of the integral can be evaluated analytically
\begin{align}
  \text{PDF}(u) & = \int_0^\infty r' dr'
  I_\text{src}(v) 
  \int_0^{2\pi} d\phi' 
  \text{PSF}(\sqrt{2\sigma^2(u+v-2\sqrt{uv}\cos(\phi-\phi'))})
  \\
  & = \int_0^\infty dv
  I_\text{src}(v) 
  \left(\frac{\gamma-1}{\gamma}\right)
  \left( \frac{\gamma}{\gamma + u + v}\right)^\gamma 
  \times {}_2F_1 \left(\gamma/2,\frac{1+\gamma}{2},1,\frac{4uv}{(\gamma+u+v)^2}\right).
\end{align}
where $v=(r'/\sigma)^2/2$ and ${}_2F_1$ is the Gaussian hypergeometric
function.  This convolution formula reduces the expected photon
distribution to a single numerical integral.

@@

For smaller and fainter extended sources, the differences in the PDF for
different spatial models are lost in the noise and the LAT is not sensitive
to the detailed spatial structure of these sources.  Therefore, in our
search for extended sources we use only a radially-symmetric uniform
surface brightness profile as our spatial extension shape. Unless otherwise noted,
we quote the radius of the edge ($\sigma$) as the size of the source.

\section{Validation of Analysis Method}

@@

We tested the false detection probability for the extended source test
by fitting the extension of point-like sources.  \cite{mattox_egret}
discuss that the test statistic distribution for a likelihood ratio test
on the existence of a source at a given position is
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\onehalf(\chi^2_1(\ts)+\delta(\ts)).
\end{equation}
The particular form of Equation \ref{ts_ext_distribution} is due to the
null hypothesis (source flux $\Phi=0$) residing on the edge of parameter
space and the model hypothesis adding a single degree of freedom and
leads to the often quoted result $\sqrt{TS}=\sigma$. It is plausible
to expect a similar distribution of the test statistic in the test for
source extension since the same conditions apply (with the source flux
$\Phi$ replaced by the source radius $r$ and $r<0$ being unphysical).
To verify Equation~\ref{ts_ext_distribution}, we evaluated the
empirical distribution function of \tsext computed from simulated sources.

We simulated point-like sources with various spectral forms using
the LAT on-orbit simulation tool
\gtobssim\footnote{\gtobssim is distributed publicly by the FSSC.} and fit the sources
with \pointlike using both point-like
and extended source hypotheses.  These point-like sources were simulated with a power-law
spectral model with integrated fluxes above 100 \mev ranging from $3\times10^{-9}$ 
to $1\times10^{-6}$ \phflux in six discrete steps and spectral
indices ranging from 1.5 to 3 in four discrete steps.  These values
were picked to represent typical source parameters of LAT-detected
sources. The point-like sources were simulated on top of an isotropic
background with a power-law spectral model with
integrated flux above 100 \mev of $1.5\times10^{-5}$ \phflux
and spectral index 2.1.
This was
taken to be the same as the isotropic spectrum measured by EGRET
\citep{sreekumar_isotropic}.  The Monte Carlo simulation was performed
over a one-year observation period using a representative 
spacecraft orbit and livetime.
The reconstruction was performed
using the P7\_V6 Source class event selection and IRFs \citep{lat_on_orbit_psf}. For each 
significantly detected point-like source ($\ts\ge25$), we used \pointlike
to fit the source as an extended source and calculate \tsext.
\newtext{
This entire procedure was performed twice, once fitting in the 1 \gev
to 100 \gev energy range and once fitting in the 10 \gev to 100 \gev
energy range.

For each set of spectral parameters, $\sim20,000$ statistically independent
simulations were performed. For dimmer spectral models, many of the
simulations left the source insignificant ($\ts<25$)
and were discarded.  Table~\ref{ts_ext_num_sims}
shows the different spectral models used in our study as well as the
number of simulations and the average point-like source
significance.  The cumulative density of \tsext is plotted in
Figure~\ref{ts_ext_mc_1000} and Figure~\ref{ts_ext_mc_10000} 
and compared to the $\chi^2_1/2$ distribution of
Equation~\ref{ts_ext_distribution}.

Our study shows broad agreement between simulations and
Equation~\ref{ts_ext_distribution}. To the extent that there is
disagreement, the simulations tended to produce smaller than expected
values of \tsext which would make the formal significance conservative.
Considering the distribution in Figure~\ref{ts_ext_mc_1000} and
Figure~\ref{ts_ext_mc_10000}, the choice of a threshold \tsext set to 16
(correspond to a formal $4\sigma$ significance) is reasonable.
}

@@

We calculated the LAT flux threshold to detect that a spatially extended
source is extended. We define the detection threshold as the flux at
which the value of $\tsext$ averaged over many statistical realizations is
$\langle\tsext\rangle=16$ for a source of a given flux and true extension
\newtext{
(corresponding to a formal $4\sigma$ significance).
}

@@

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extensions varying from $\sigma=0\fdg1$
to $2\fdg0$.  
The threshold is high for small extensions when the
source is 
small compared to the size of the PSF. 
It drops quickly with increasing source size and reaches
a minimum around 0\fdg5. 
The threshold increases
for large extended sources because the source becomes
increasingly diluted by the background.
Figure~\ref{index_sensitivity} shows
the threshold using photons with energies between 100 \mev and 100 \gev
and also using only photons with energies between 1 \gev and 100 \gev.
Except for very large or very soft
sources, our detection threshold is
not substantially improved by including photons with energies between 100 \mev and
1 \gev.  This is also demonstrated in Figure~\ref{four_plots_ic443}
which shows \tsext for the SNR IC443 computed independently in twelve
energy bins between 100 \mev and 100 \gev. For IC443, which has a
spectral index $\sim2.4$ and an extension $\sigma=\sim0\fdg35$, 
almost the entire increase in likelihood
modeling the source as being extended comes
from energies above 1 \gev.  Furthermore, other systematic errors
become increasingly important at low energy. For our extension search
(Section~\ref{extended_source_search_method}),
we therefore used only photons with energies above 1 \gev.

@@

It is impossible to discriminate using only LAT data between a
spatially extended source and multiple point-like sources separated by
angular distances comparable to or smaller than the size of the LAT PSF.
To assess the plausibility of source confusion for situations
in which $\tsext>16$, we developed an algorithm in \pointlike to
simultaneously fit the position and spectrum of two point-like sources.

After simultaneously fitting the two positions and two spectra,
we define \tsinc as twice the increase in the log of the likelihood
fitting the region with two point-like sources compared to fitting the
region with one point-like source:
\begin{equation}
  \tsinc=2\log(\likelihood_\text{2pts}/\likelihood_\text{ps}).
\end{equation} 
For the following analysis of LAT data, \tsinc was computed
by fitting the spectra of the two point-like sources in \gtlike using the best fit positions
of the sources found by \pointlike.

@@

We did this study first in the energy range from 1 \gev to 100 \gev by
simulating extended sources of flux $4\times10^{-9}$ \phflux integrated
from 1 \gev to 100 \gev with a power-law spectral model with
spectral index 2.  This flux was picked to be representative of the
new extended sources that were discovered in the following analysis
when looking in the 1 \gev to 100 \gev energy range
(see Section~\ref{new_ext_srcs_section}).
We simulated these sources using a uniform surface brightness profile
with extensions varying
up to $1\degree$.  
\newtext{
Figure~\ref{confusion_plot}a shows 
the distribution of $\tsext-\tsinc$ as a
function of the simulated extension of the source
for 200 statistically independent simulations.
}

Figure~\ref{confusion_plot}c shows the same plot but when fitting
simulations of two point-like sources each with half of the flux of
the spatially extended source and with the same spectral index as the
extended source.  Finally, figure~\ref{confusion_plot}e shows the same
plot with each point-like source having the same flux but different
spectral indices.  One point-like source had a spectral index of 1.5
and the other an index of 2.5.

The same three plots are shown in figure~\ref{confusion_plot}b,
figure~\ref{confusion_plot}d, and figure~\ref{confusion_plot}f but this
time when analyzing a source of flux $10^{-9} \phflux$ (integrated from
10 \gev to 100 \gev) only in the 10 \gev to 100 \gev energy range.
This spectrum is typical of the new extended sources discovered
using only photons with energies between 10 \gev and 100 \gev (see
Section~\ref{new_ext_srcs_section}).

Several interesting conclusions can be made from this study. The overall
behavior of these plots is what one would expect. \tsext is on average
greater than \tsinc when fitting simulated extended sources whereas
\tsinc is greater than \tsext when fitting two simulated point-like sources.
Furthermore, the distributions of \tsext and \tsinc separate better
when the extended source is bigger or when the two point-like sources are
farther apart.

The distributions also show that for simulated
extended sources with typical
sizes ($\sigma\sim0\fdg5$), one can often obtain almost as large
an increase in likelihood
with two sources as with one extended source ($\tsinc\sim\tsext$).
This is because although
two point-like sources represent an incorrect spatial model, 
the two point-like sources have four additional degrees
of freedom (two spatial and two spectral parameters)
and can therefore easy model
the extended source and statistical fluctuations in the data.
This effect is most pronounced when using photons with energies between 1
\gev and 100 \gev where the PSF is larger.  
On the other hand,
when testing the two simulated point-like sources for extension our test
found that \tsinc was larger than \tsext.
This is because the one additional extension parameter is less able to
model the second source or statistical fluctuations.
From this, we conclude that the $\tsext>\tsinc$ test is 
powerful at discarding cases in which the true emission is derived
from two point-like sources instead of from a spatially extended
source.

From this Monte Carlo study, we can see the limits of a LAT analysis
of spatially extended sources.  Section~\ref{monte_carlo_validation}
showed that we have a statistical test that finds when a LAT source is
not well described by the PSF.  But this test does not uniquely prove
that the emission originates from spatially extended emission instead
of from multiple unresolved sources.  Demanding that $\tsext>\tsinc$
is a powerful second test to avoid cases of simple confusion of two
point-like sources. But it could always be the case that an extended
source candidate is actually the superposition of multiple point-like or
extended sources that could be resolved with deeper observations of the
region.  There is nothing about this conclusion unique to LAT analysis,
but the broad PSF of the LAT and the density of sources expected to be
\gev emitters in the galactic plane makes this issue more significant
for LAT analysis.  When possible, multiwavelength information should be
used to help select the best model of the sky.

@@

For all following LAT analysis, we used the same two-year dataset
that was used in 2FGL spanning from 2008 August 4 to 2010 August 1. We
applied the same acceptance cuts and we used the same P7\_V6 Source class
event selection and IRFs \citep{lat_on_orbit_psf}.  
When analyzing sources in \pointlike, we used a circular $10\degree$ region of
interest (ROI) centered on our source.  
When refitting the region in \gtlike using the best fit spatial and
spectral models from \pointlike, we used the `binned likelihood' mode of
\gtlike on a $14\degree\times14\degree$ ROI with a pixel size of 0\fdg03.

@@

As further validation of our method for studying
spatially extended sources, we reanalyzed the twelve spatially extended
sources which were included in 2FGL \citep{second_cat}.  Even though
each of these sources required a dedicated analysis and they
were fit with a variety of spatial models, it is valuable to show that
these sources are significantly extended using our systematic 
method that assumed a radially-symmetric uniform surface
brightness profile.  On the other hand, for some of
these sources a uniform disk spatial model does not well describe the
observed extended emission and so the dedicated LAT publications provides
a qualitatively better model of these sources.

@@

% 3 PWN
Three PWNe, MSH\,15$-$52, Vela X, and HESS\,J1825$-$137 were
identified in 2FGL \citep{msh1552,velax,fermi_hess_j1825}.  
HESS\,J1825$-$137 was significantly detected using photons
with energies between 10 \gev and 100 \gev.
To improve our analysis of this source, we had to remove
from our background model 
2FGL\,J1823.1$-$1338c which was included in 2FGL to
represent part of the extended emission.
To avoid confusion with the nearby bright pulsar PSR\,J1509$-$5850, MSH\,15$-$52 had
to be
analyzed at high energies.  Using photons with energies above 10 \gev,
we fit the extension of MSH\,15$-$52 to be consistent with the published
size but with \tsext=6.6.

@@

We estimated a second systematic error on the extension of a source
due to uncertainty in our model of the Galactic diffuse emission by
using an alternative model which takes as input templates
calculated by
GALPROP\footnote{GALPROP is a software package for calculating the
Galactic $\gamma$-ray emission based on a model of cosmic-ray propagation
in the Galaxy \citep{galprop1998,galprop2011}. 
See also \url{http://galprop.stanford.edu/} for details.} 
but then fits each template locally in
the surrounding region.
The particular GALPROP model that was used as input is described in
the LAT analysis of the isotropic diffuse
emission \citep{isotropic_lat}.  
The intensities of various components
of the Galactic diffuse emission were fitted individually using a
spatial distribution of the intensities within the ROI as predicted by
the model.  We distinguished contributions from cosmic-ray interactions with the
molecular hydrogen, the atomic and ionized hydrogen, residual gas traced
by dust \citep{isabelle_dark_gass}, and the interstellar radiation
field. We further split the contributions from interactions with molecular
and atomic hydrogen to the Galactic diffuse emission according to the
distance from the Galactic center in which they are produced. Hence, we
replaced the standard diffuse emission model by 18 individually fitted
templates to describe individual components of the diffuse emission.
A similar crosscheck was used in an analysis of RX\,J1713.7$-$3946 
by the LAT Collaboration \citep{rx_j1713_lat}.

@@

\begin{deluxetable}{rrrrrrrrrrrrrrrrrrrrrrr}

>>>>

The current generation of air Cherenkov detectors have
made it apparent that many sources can be spatially resolved
at even higher energies.
Most
prominent was a survey of the Galactic plane using the High Energy
Stereoscopic System (H.E.S.S) which reported 14 spatially extended
sources with extensions varying from $\sim0\fdg1$ to $\sim0\fdg25$
\citep{hess_plane_survey}.  In fact, within our Galaxy only very few
sources detected at \tev energies, most notably the $\gamma$-ray binaries
LS\,5039 \citep{HESSLS5039}, LS I+61$-$303 \citep{MAGICLSI, VERITASLSI},
HESS\,J0632+057 \citep{HESS0632}, and the Crab nebula \citep{crab_weekes},
have no detectable extension.  High-energy $\gamma$-rays from
\tev sources are produced by the decay of $\pi^0$s produced by hadronic
interactions with interstellar matter and by relativistic electrons
due to Inverse Compton (IC) scattering and bremsstrahlung radiation.  
It is plausible that the \gev and
\tev emission from these sources originates from the same population of
high-energy particles and so at least some of these sources should be
detectable at \gev energies.  Studying 
these \tev sources at \gev energies would help to
determine the emission mechanisms producing these high energy photons.

@@

\newtext{
A new maximum-likelihood analysis tool has been developed to address the
unique requirements for studying spatially extended sources with the LAT.
It works by maximizing the Poisson likelihood to find the observing counts
given a parametrized spatial and spectral model of the sky.  
Space is
divided into cubes of position and energy using the HEALPix representation of
the sky \citep{healpix} and the likelihood is maximized over all bins in
a region.
}
The extension of a source can be modeled by a geometric shape
(e.g. a disk or a two-dimensional Gaussian) and the position, extension,
and spectrum of the source can be simultaneously fit.

This type of analysis is not feasible using the standard LAT likelihood
analysis tool \gtlike\footnote{\gtlike is distributed publicly by the
FSSC.} because it can only fit the spectral parameters of the model
unless a more sophisticated iterative procedure is used.  
We note that \gtlike has been used in the
past in several studies of source extension in the LAT Collaboration
\citep{lmc,smc,w28,w51c}.  In these studies, a profile based upon a
set of \gtlike maximum likelihood fits at fixed extensions was used
to build a profile of the likelihood as a function of extension.
\newtext{
The \gtlike likelihood profile approach has been show to correctly
reproduce the extension of simulated extended sources assuming that the
true position is known \citep{francesco_2011}.  But it is not optimal
because the position, extension, and spectrum of the source must be
simultaneously fit to find the best fit parameters and to maximize the
statistical significance of the detection.  Furthermore, because the \gtlike
approach is computationally intensive, no large-scale Monte Carlo
simulations have been run to calculate its false detection rate.
}

@@

In \pointlike, it is assumed that the spatial and spectral model
of an extended source are separable, i.e. that the source model
$M(l,b,E)=S(l,b)\times X(E)$ where $S(l,b)$ is the spatial distribution
and $X(E)$ is the spectral distribution.  To fit an extended source,
\pointlike convolves the extended source shape with the PSF (as a function
of energy) and uses the \minuit library \citep{minuit_documentation}
to maximize the likelihood by simultaneously varying the position,
extension, and spectrum of the source.  As will be described in
Section~\ref{monte_carlo_validation}, simultaneously fitting the
position, extension, and spectrum is important to maximize
the statistical significance of the detection of the extension of a source.
To avoid projection effects, the longitude and latitude 
of the source are
not directly fit but instead the displacement of the source in a rotated
reference frame.

@@

For extended sources with an assumed radially-symmetric shape,
we optimized the calculation by performing one
of the integrals analytically.
The expected photon 
distribution can be written as
\begin{equation}
  \text{PDF}(\vec r) = \int  \text{PSF}(|\vec r - \vec r'|)I_\text{src}(\vec r') r' dr' d\phi'
\end{equation}
where $I_\text{src}(\vec r')$ is the spatial distribution of the
source.
\newtext{
The PSF of the LAT is currently parameterized 
in the Pass~7\_V6 (P7\_V6) Source Instrument
Response Function \citep[IRFs,][]{lat_on_orbit_psf} by a King function \citep{king_function}:
}
\begin{equation}
  \text{PSF}(r) = 
  \frac{1}{2\pi\sigma^2}
  \left(1-\frac{1}{\gamma}\right)
  \left(1+\frac{u}{\gamma}\right)^{-\gamma},
\end{equation}
where $u=(r/\sigma)^2/2$ and $\sigma$ and $\gamma$ are free parameters
\citep{matthew_kerr_thesis}.  For radially-symmetric extended sources,
the angular part of the integral can be evaluated analytically
\begin{align}
  \text{PDF}(u) & = \int_0^\infty r' dr'
  I_\text{src}(v) 
  \int_0^{2\pi} d\phi' 
  \text{PSF}(\sqrt{2\sigma^2(u+v-2\sqrt{uv}\cos(\phi-\phi'))})
  \\
  & = \int_0^\infty dv
  I_\text{src}(v) 
  \left(\frac{\gamma-1}{\gamma}\right)
  \left( \frac{\gamma}{\gamma + u + v}\right)^\gamma 
  \times {}_2F_1 \left(\gamma/2,\frac{1+\gamma}{2},1,\frac{4uv}{(\gamma+u+v)^2}\right).
\end{align}
where $v=(r'/\sigma)^2/2$ and ${}_2F_1$ is the Gaussian hypergeometric
function.  This convolution formula reduces the expected photon
distribution to a single numerical integral.

@@

For smaller and fainter extended sources, the differences in the PDF for
different spatial models are lost in the noise and the LAT is not sensitive
to the detailed spatial structure of these sources.  Therefore, in our
search for extended sources we use only a radially-symmetric uniform
surface brightness profile as our spatial extension shape. Unless otherwise noted,
we quote the radius to the edge ($\sigma$) as the size of the source.

\section{Validation of the Analysis Method}

@@

We tested the false detection probability for the extended source test
by fitting the extension of point-like sources.  \cite{mattox_egret}
discuss that the test statistic distribution for a likelihood ratio test
on the existence of a source at a given position is
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\onehalf(\chi^2_1(\ts)+\delta(\ts)).
\end{equation}
The particular form of Equation \ref{ts_ext_distribution} is due to the
null hypothesis (source flux $\Phi=0$) residing on the edge of parameter
space and the model hypothesis adding a single degree of freedom.
It leads to the often quoted result $\sqrt{TS}=\sigma$. It is plausible
to expect a similar distribution for the test statistic in the test for
source extension since the same conditions apply (with the source flux
$\Phi$ replaced by the source radius $r$ and $r<0$ being unphysical).
To verify Equation~\ref{ts_ext_distribution}, we evaluated the
empirical distribution function of \tsext computed from simulated sources.

We simulated point-like sources with various spectral forms using
the LAT on-orbit simulation tool
\gtobssim\footnote{\gtobssim is distributed publicly by the FSSC.} and fit the sources
with \pointlike using both point-like
and extended source hypotheses.  These point-like sources were simulated with a power-law
spectral model with integrated fluxes above 100 \mev ranging from $3\times10^{-9}$ 
to $1\times10^{-5}$ \phflux and spectral
indices ranging from 1.5 to 3.  These values
were picked to represent typical parameters of LAT-detected
sources. The point-like sources were simulated on top of an isotropic
background with a power-law spectral model with
integrated flux above 100 \mev of $1.5\times10^{-5}$ \phflux
and spectral index 2.1.
This was
taken to be the same as the isotropic spectrum measured by EGRET
\citep{sreekumar_isotropic}.  The Monte Carlo simulation was performed
over a one-year observation period using a representative 
spacecraft orbit and livetime.
The reconstruction was performed
using the P7\_V6 Source class event selection and IRFs \citep{lat_on_orbit_psf}. For each 
significantly detected point-like source ($\ts\ge25$), we used \pointlike
to fit the source as an extended source and calculate \tsext.
\newtext{
This entire procedure was performed twice, once fitting in the 1 \gev
to 100 \gev energy range and once fitting in the 10 \gev to 100 \gev
energy range.

For each set of spectral parameters, $\sim20,000$ statistically independent
simulations were performed. For dimmer spectral models, many of the
simulations left the source insignificant ($\ts<25$)
and were discarded.  Table~\ref{ts_ext_num_sims}
shows the different spectral models used in our study as well as the
number of simulations and the average point-like source
significance.  The cumulative density of \tsext is plotted in
Figure~\ref{ts_ext_mc_1000} and \ref{ts_ext_mc_10000} 
and compared to the $\chi^2_1/2$ distribution of
Equation~\ref{ts_ext_distribution}.

Our study shows broad agreement between simulations and
Equation~\ref{ts_ext_distribution}. To the extent that there is
a discrepancy, the simulations tended to produce smaller than expected
values of \tsext which would make the formal significance conservative.
Considering the distribution in Figure~\ref{ts_ext_mc_1000} and
Figure~\ref{ts_ext_mc_10000}, the choice of a threshold \tsext set to 16
(correspond to a formal $4\sigma$ significance) is reasonable.
}

@@

We calculated the LAT flux threshold to detect that a spatially extended
source is extended. We define the detection threshold as the flux at
which the value of $\tsext$ averaged over many statistical realizations is
$\langle\tsext\rangle=16$ 
\newtext{
(corresponding to a formal $4\sigma$ significance)
for a source of a given flux and extension.
}

@@

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extensions varying from $\sigma=0\fdg1$
to $2\fdg0$.  
The threshold is high for small extensions when the
source is 
small compared to the size of the PSF. 
It drops quickly with increasing source size and reaches
a minimum around 0\fdg5. 
The threshold increases
for large extended sources because the source becomes
increasingly diluted by the background.
Figure~\ref{index_sensitivity} shows
the threshold using photons with energies between 100 \mev and 100 \gev
and also using only photons with energies between 1 \gev and 100 \gev.
Except for very large or very soft
sources, our detection threshold is
not substantially improved by including photons with energies between 100 \mev and
1 \gev.  This is also demonstrated in Figure~\ref{four_plots_ic443}
which shows \tsext for the SNR IC443 computed independently in twelve
energy bins between 100 \mev and 100 \gev. For IC443, which has a
spectral index $\sim2.4$ and an extension $\sim0\fdg35$, 
almost the entire increase in likelihood
modeling the source as being extended comes
from energies above 1 \gev.  Furthermore, other systematic errors
become increasingly large at low energy. For our extension search
(Section~\ref{extended_source_search_method}),
we therefore used only photons with energies above 1 \gev.

@@

It is impossible to discriminate using only LAT data between a
spatially extended source and multiple point-like sources separated by
angular distances comparable to or smaller than the size of the LAT PSF.
To assess the plausibility of source confusion for situations
in which $\tsext\ge16$, we developed an algorithm in \pointlike to
simultaneously fit the position and spectrum of two point-like sources.

After simultaneously fitting the two positions and two spectra,
we define \tsinc as twice the increase in the log of the likelihood
fitting the region with two point-like sources compared to fitting the
region with one point-like source:
\begin{equation}
  \tsinc=2\log(\likelihood_\text{2pts}/\likelihood_\text{ps}).
\end{equation} 
\newtext{
For the following analysis of LAT data, \tsinc was computed
by fitting the spectra of the two point-like sources in \gtlike using the best fit positions
of the sources found by \pointlike.
}

@@

We did this study first in the energy range from 1 \gev to 100 \gev by
simulating extended sources of flux $4\times10^{-9}$ \phflux integrated
from 1 \gev to 100 \gev and a power-law spectral model with
spectral index 2.  This flux was picked to be representative of the
new extended sources that were discovered in the following analysis
when looking in the 1 \gev to 100 \gev energy range
(see Section~\ref{new_ext_srcs_section}).
We simulated these sources using a uniform surface brightness profile
with extensions varying
up to $1\degree$.  
\newtext{
Figure~\ref{confusion_extended_plot}a shows the distribution
of \tsext and \tsinc and 
Figure~\ref{confusion_extended_plot}c shows 
the distribution of $\tsext-\tsinc$ as a
function of the simulated extension of the source
for 200 statistically independent simulations.

Figure~\ref{confusion_2pts_plot}a shows the same plot but when fitting
two simulated point-like sources each with half of the flux of
the spatially extended source and with the same spectral index as the
extended source.  Finally, figure~\ref{confusion_2pts_plot}c shows the same
plot with each point-like source having the same flux but different
spectral indices.  One point-like source had a spectral index of 1.5
and the other an index of 2.5.

The same four plots are shown in figure~\ref{confusion_extended_plot}b, \ref{confusion_extended_plot}d, 
\ref{confusion_2pts_plot}b, and \ref{confusion_2pts_plot}d but this
time when analyzing a source of flux $10^{-9}$ \phflux (integrated from
10 \gev to 100 \gev) only in the 10 \gev to 100 \gev energy range.
This spectrum is typical of the new extended sources discovered
using only photons with energies between 10 \gev and 100 \gev (see
Section~\ref{new_ext_srcs_section}).

Several interesting conclusions can be made from this study. The overall
behavior of these plots is what one would expect. $\tsext-\tsinc$
is mostly positive when fitting the simulated extended sources in
Figure~\ref{confusion_extended_plot} whereas $\tsext-\tsinc$ is
mostly negative when fitting the two simulated point-like sources in
Figure~\ref{confusion_2pts_plot}.  
Furthermore, $\tsext-\tsinc$ becomes more positive when the extended
soruce becomes larger and more negative when the two point-like sources
are farther apart.
From this, we conclude that the $\tsext>\tsinc$ test is
powerful at discarding cases in which the true emission comes from
two point-like sources without frequently rejecting real extended sources.

The other interesting feature in Figure~\ref{confusion_extended_plot}a
and \ref{confusion_extended_plot}b is that for simulated extended
sources with typical sizes ($\sigma\sim0\fdg5$), one can often obtain
almost as large an increase in likelihood fitting the source as two
point-like sources ($\tsinc\sim\tsext$).  This is because although the
two point-like sources represent an incorrect spatial model, the second
source has four additional degrees of freedom (two spatial and two
spectral parameters) and can therefore easy model much of the extended
source and statistical fluctuations in the data.  This effect is most
pronounced when using photons with energies between 1 \gev and 100 \gev
where the PSF is broader.
}

From this Monte Carlo study, we can see the limits of a LAT analysis
of spatially extended sources.  Section~\ref{monte_carlo_validation}
showed that we have a statistical test that finds when a LAT source is
not well described by the PSF.  But this test does not uniquely prove
that the emission originates from spatially extended emission instead
of from multiple unresolved sources.  Demanding that $\tsext>\tsinc$
is a powerful second test to avoid cases of simple confusion of two
point-like sources. But it could always be the case that an extended
source is actually the superposition of multiple point-like or
extended sources that could be resolved with deeper observations of the
region.  There is nothing about this conclusion unique to LAT analysis,
but the broad PSF of the LAT and the density of sources expected to be
\gev emitters in the galactic plane makes this issue more significant
for LAT analysis.  When possible, multiwavelength information should be
used to help select the best model of the sky.

@@

For all following LAT analysis, we used the same two-year dataset
that was used in 2FGL spanning from 2008 August 4 to 2010 August 1. We
applied the same acceptance cuts and we used the same P7\_V6 Source class
event selection and IRFs \citep{lat_on_orbit_psf}.  
When analyzing sources in \pointlike, we used a circular $10\degree$ region of
interest (ROI) centered on our source and eight energy bins per
logarithmic decade in energy.
When refitting the region in \gtlike using the best fit spatial and
spectral models from \pointlike, we used the `binned likelihood' mode of
\gtlike on a $14\degree\times14\degree$ ROI with a pixel size of 0\fdg03.

@@

As further validation of our method for studying
spatially extended sources, we reanalyzed the twelve spatially extended
sources which were included in 2FGL \citep{second_cat}.  Even though
each of these sources required a dedicated analysis and they
were fit with a variety of spatial models, it is valuable to show that
these sources are significantly extended using our systematic 
method assuming a radially-symmetric uniform surface
brightness profile.  On the other hand, for some of
these sources a uniform disk spatial model does not well describe the
observed extended emission and so the dedicated LAT publications provides
a qualitatively better model of these sources.

@@

% 3 PWN
Three PWNe, MSH\,15$-$52, Vela X, and HESS\,J1825$-$137, were
identified in 2FGL \citep{msh1552,velax,fermi_hess_j1825}.  
HESS\,J1825$-$137 was significantly detected using photons
with energies between 10 \gev and 100 \gev.
To improve our analysis of this source, removed
from our background model 
2FGL\,J1823.1$-$1338c which was included in 2FGL to
represent part of the extended emission.
To avoid confusion with the nearby bright pulsar PSR\,J1509$-$5850, MSH\,15$-$52 had
to be
analyzed at high energies.  Using photons with energies above 10 \gev,
we fit the extension of MSH\,15$-$52 to be consistent with the published
size but with \tsext=6.6.

@@

We estimated a second systematic error on the extension of a source
due to uncertainty in our model of the Galactic diffuse emission by
using an alternative model which takes as input templates
calculated by
GALPROP\footnote{GALPROP is a software package for calculating the
Galactic $\gamma$-ray emission based on a model of cosmic-ray propagation
in the Galaxy \citep{galprop1998,galprop2011}. 
See also \url{http://galprop.stanford.edu/} for details.} 
but then fits each template locally in
the surrounding region.
The particular GALPROP model that was used as input is described in
the LAT analysis of the isotropic diffuse
emission \citep{isotropic_lat}.  
The intensities of various components
of the Galactic diffuse emission were fitted individually using a
spatial distribution predicted by
the model.  We distinguished contributions from cosmic-ray interactions with the
molecular hydrogen, the atomic and ionized hydrogen, residual gas traced
by dust \citep{isabelle_dark_gass}, and the interstellar radiation
field. We further split the contributions from interactions with molecular
and atomic hydrogen to the Galactic diffuse emission according to the
distance from the Galactic center in which they are produced. Hence, we
replaced the standard diffuse emission model by 18 individually fitted
templates to describe individual components of the diffuse emission.
A similar crosscheck was used in an analysis of RX\,J1713.7$-$3946 
by the LAT Collaboration \citep{rx_j1713_lat}.

@@

\begin{deluxetable}{rr|rrrrrrrrrrrrrrrrrrrrr}
